---
title: "Day 1"
author: "john"
date: "12/8/2019"
output: github_document
---

```{r}
library(advent)
```


# Introduction

I found out about Advent of Code this year and I really love the idea! I will be documenting my progress here. I hope you enjoy reading my thoughts.

# Utilities

First I need to write a few utilities for getting data from the Advent of Code website and copying it to my clipboard.

## Writing to Clipboard

I want to copy my answers to the clipboard so I can paste them into the website. Here I have a function to detect the operating system, then copy to the clipboard if it is MacOS or Windows.

```{r}
is_mac
```

```{r}
is_windows
```

```{r}
write_to_clipboard
```

Note that `write_to_clipboard` invisibly returns its argument. This allows me to put it into a chain of composed functions and keep the chain evaluating.

## Get Data from Advent of Code

I want to download the input data from Advent of Code, but I need to pass along my session cookie. I do this by using the httr package, storing the session cookie in an environment variable in my .Rsession file.

```{r}
get_data
```


